Testable normative theories

### 2020-05-25 simple.tex
Started using sequences (history) instead of pairs of args

\subsection{Definition}
A normative theory is a tuple $(\allprops, ¬, \allargs, I, (\ileadstoe[i])_{i \in I})$, 
with ${\ileadstoe[i]} \subseteq (\allargs × \allargs) × \Phi$, 
where $\Phi = \allprops × \set{\text{possible}, \text{sure}}$.

The symbol $¬$ is a negation operator on $\allprops$, such that $\forall \prop \in \allprops: ¬¬\prop = \prop$ (this article also uses this symbol to negate classical logic propositions and as a negation operator on $\Phi$, which should create no confusion). Define $¬$ on $\Phi$ as follows, $\forall \prop \in \allprops$: $¬(\prop, \text{possible}) = (¬\prop, \text{sure})$ and $¬(\prop, \text{sure}) = (¬\prop, \text{possible})$. Observe that $¬¬\phi = \phi$.

Given any $\ar', \ar, \prop$, let $(\ar', \ar) \ileadstost (\prop, \text{sure})$ be equivalent to $(\ar', \ar) \nileadstoe (¬\prop, \text{poss}) \land (\ar', \ar) \ileadstoe (\prop, \text{sure}) \land (\ar', \ar) \nileadstoe (¬\prop, \text{sure})$; and $(\ar', \ar) \ileadstost (\prop, \text{poss})$ be equivalent to $(\ar', \ar) \nileadstoe (¬\prop, \text{sure})$.
\footnote{Assume $i$, given $\ar$ and $\ar'$, picks repeatedly (in various circumstances) nonempty subsets of $\set{\prop, ¬\prop}$. Define $(\ar', \ar) \ileadstoe (\prop, \text{sure})$ iff she sometimes chooses $\set{\prop}$ and $(\ar', \ar) \ileadstoe (\prop, \text{poss})$ iff she sometimes chooses $\set{\prop}$ or $\set{\prop, ¬\prop}$. Then, $(\ar', \ar) \ileadstost (\prop, \text{sure})$ iff she always chooses $\set{\prop}$ and $(\ar', \ar) \ileadstost (\prop, \text{poss})$ iff she always chooses either $\set{\prop}$ or $\set{\prop, ¬\prop}$.}

To lighten notations, $\ileadstoe[i]$ will generally be denoted by $\ileadstoe$, this symbol being understood as the result of a function of $i$. When the symbol $i$, $\ileadstoe$ or $\ileadstost$ is used in a formula without $i$ being bound, the binding $\forall i \in I$ is implied.

As usual with binary relations, $(\ar', \ar) \nileadstost \phi$ means $¬[(\ar', \ar) \ileadstost \phi]$.
Abusing notation, given $\ar \in \allargs$ and $\phi \in \Phi$, I write $\ar \ileadstost \phi$ to mean $[\forall \ar' \in \allargs: (\ar', \ar) \ileadstost \phi]$.

\subsection{Classify propositions}
Define $(\ar_1, \ar) \ileadstostp \phi$ iff $(\ar_1, \ar) \nileadstoe ¬\phi \land (\ar_1, \ar) \ileadstoe \phi$. Define $\ar \ileadstostp \phi$ iff $\forall \ar_1 \in \allargs: (\ar_1, \ar) \ileadstostp \phi$.

\begin{axiom}[No contradiction]
	\label{ax:nocontr}
	$\forall \phi \in \Phi: \phi \in \iPhi ⇒ ¬\phi \notin \iPhi$.
\end{axiom}
\begin{axiom}[Normative adequacy]
	\label{ax:norm}
	$\forall \prop \in \allprops$: 
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstostp \propposs] ⇒ \propposs \in \iPhi;
	\end{equation}
	and
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstostp \propsure \land \ar \ileadstostp \propposs] ⇒ \propsure \in \iPhi.
	\end{equation}
\end{axiom}

@@

\section{Empirical theories}
\subsection{Definition}
An empirical theory for a normative theory is a tuple $(q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, +)$,
with ${\gleadsto[q(i)]} \subseteq \allargs × \Phi$, ${\gbeats} \subseteq \allargs × \allargs$,
and $+$ a commutative associative internal binary operation on $\allargs$. Given a finite $\args \subseteq \allargs$, the notation $\sum \args = \sum_{\ar \in \args} \ar$ thus designates a well-defined argument. 
\commentOC{Check about finiteness.}

@@

	To say that $\phi \in \iPhi$, need to not depend on choice of $q$, in the following sense: if with $q$, $\phi$ seems stable, but with $q'$, $¬\phi$ seems stable, then $\phi$ is not proved as being in $\iPhi$. But if with $q$, $\phi$ seems stable, but with $q'$, nothing can be shown to be stable, then $\phi$ is proved (temporarily) as being in $\iPhi$. More precisely, after $q'$, the model must be able to stabilize $i$ with its $q$.
\end{remark}

>>>>

\subsection{Definition}
A normative theory is a tuple $(\allprops, ¬, \allargs, I, (\ileadstoe[i])_{i \in I})$, 
with ${\ileadstoe[i]} \subseteq \allhist × \Phi$, 
where $\allhist$ denotes all finite sequences of elements in $\allargs$ (including the empty sequence) and $\Phi = \allprops × \set{\text{possible}, \text{sure}}$.

The symbol $¬$ is a negation operator on $\allprops$, such that $\forall \prop \in \allprops: ¬¬\prop = \prop$. This article also uses this symbol as a negation operator on $\Phi$, defined as follows, $\forall \prop \in \allprops$: $¬(\prop, \text{possible}) = (¬\prop, \text{sure})$ and $¬(\prop, \text{sure}) = (¬\prop, \text{possible})$. Observe that $¬¬\phi = \phi$. (The symbol $¬$ is also used in this article to negate classical logic propositions; this should create no confusion.)
As usual with binary relations, $\hist \nileadstoe[i] \phi$ means $(\hist, \phi) \notin {\ileadstoe[i]}$.

Given any $\ar_0 \in \allargs, \phi \in \Phi$, define $\ar_0 \ileadstost[i] \phi$ iff $\forall \hist \in \allhist, \ar_1 \in \allargs: (\hist, \ar_0) \nileadstoe[i] ¬\phi \land (\hist, \ar_0) \ileadstoe[i] \phi \land (\hist, \ar_0, \ar_1) \nileadstoe[i] ¬\phi \land (\hist, \ar_0, \ar_1) \ileadstoe[i] \phi$.
\footnote{TODO Assume $i$, given $\ar$ and $\ar'$, picks repeatedly (in various circumstances) nonempty subsets of $\set{\prop, ¬\prop}$. Define $(\ar', \ar) \ileadstoe (\prop, \text{sure})$ iff she sometimes chooses $\set{\prop}$ and $(\ar', \ar) \ileadstoe (\prop, \text{poss})$ iff she sometimes chooses $\set{\prop}$ or $\set{\prop, ¬\prop}$. Then, $(\ar', \ar) \ileadstost (\prop, \text{sure})$ iff she always chooses $\set{\prop}$ and $(\ar', \ar) \ileadstost (\prop, \text{poss})$ iff she always chooses either $\set{\prop}$ or $\set{\prop, ¬\prop}$.}

To lighten notations, $\ileadstoe[i]$ and $\ileadstost[i]$ will generally be denoted by $\ileadstoe$ and $\ileadstost$, these symbols being understood as the result of a function of $i$. 
When the symbol $\ileadstoe$ or $\ileadstost$ is used in a formula without $i$ being bound, the binding $\forall i \in I$ is implied.

\subsection{Classify propositions}
\commentOC{Check if I current definition of $\ileadstost$ is correct, in which case I can get rid of $\ileadstostp$.}
\begin{axiom}[No contradiction]
	\label{ax:nocontr}
	$\forall \phi \in \Phi: \phi \in \iPhi ⇒ ¬\phi \notin \iPhi$.
\end{axiom}
\begin{axiom}[Normative adequacy]
	\label{ax:norm}
	$\forall \prop \in \allprops$: 
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstostp \propposs] ⇒ \propposs \in \iPhi;
	\end{equation}
	and
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstostp \propsure \land \ar \ileadstostp \propposs] ⇒ \propsure \in \iPhi.
	\end{equation}
\end{axiom}

@@

\section{Empirical theories}
Given $\hist \in \allhist$ and $i \in I$, the claim $\hist \ileadstoe[i] \phi$ may not be verifiable, and is too specific to be falsifiable. However, given $\hist \in \allhist$, a broader claim is falsifiable: that for a given set of individuals $Q \subseteq I, \forall i \in Q, \hist \ileadstoe[i] \phi$ (provided $Q$ is big enough). Indeed, that claim it is equivalent to $\nexists i \in Q \suchthat \hist \nileadstoe[i] \phi$; exhibiting an $i \in Q$ such that $\hist \nileadstoe[i] \phi$ thus falsifies the broader claim.

\subsection{Definition}
An empirical theory for a normative theory is a tuple $(q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, +)$,
with ${\gleadsto[q(i)]} \subseteq \allargs × \Phi$, ${\gbeats} \subseteq \allargs × \allargs$,
and $+$ a commutative associative internal binary operation on $\allargs$. Given a finite $\args \subseteq \allargs$, the notation $\sum \args = \sum_{\ar \in \args} \ar$ thus designates a well-defined argument. 
\commentOC{Check about finiteness.}

@@

	To say that $\phi \in \iPhi$, need to not depend on choice of $q$, in the following sense: if with $q$, $\phi$ seems stable, but with $q'$, $¬\phi$ seems stable, then $\phi$ is not proved as being in $\iPhi$. But if with $q$, $\phi$ seems stable, but with $q'$, nothing can be shown to be stable, then $\phi$ is proved (temporarily) as being in $\iPhi$. More precisely, after $q'$, the model must be able to stabilize $i$ with its $q$.
\end{remark}
I think I need the following. Define $\ar_1 \ibeatse \ar_0$ iff $\exists \hist \suchthat (\hist, \ar_0, \ar_1) \ileadstoe ¬\phi$. The model claims that if $\ar_1 \ibeatse \ar_0$ and $\ar_3 \ibeatse \ar_0 + \ar_2$, then $(\hist, \ar_0 + \ar_2 + \ar_4) \ileadstoe \phi$ (for some suitable $\hist$ that does not contain yet this arg or their c-a, or perhaps just claim that $\exists \hist \suchthat \hist \ileadstoe \phi$ with $\hist$ containing this arg).

Better: the model claims that if $\ar_1 \ileadstoe ¬\phi$ and if $\ar_0$ and $\ar_1$ are in $\hist$ and none of the anticipated attackers of $\ar_0 + \ar_2$ are, then $(\hist, \ar_0 + \ar_2) \ileadstoe \phi$ (or, at least, that the model has at least one available candidate).
